---
title: "Intermediate SQL"
author: "Isaac Baca"
date: "`r Sys.Date()`"
output: html_document
---

---

# Introduction

This project uses intermediate SQL queries on three separate data sets:  

- A `films` data set which includes the titles of films, the year they were released, the country, the duration in minutes, the language of the film, the certification, gross revenue, and the budget of the film.

- A `People` data set which includes the name, birthdate, and death date of famous people.

- A `Reviews` data set which includes a film ID, user number, critic number, imdb score, number of votes, and number of facebook likes.

- For brevity of output, `limit 10` has been added to queries that contain large amounts of data.

---

# Table of contents:

1. [View the tables](#one)

2. [Selecting and Counting Data](#two)

3. [Filtering Numbers](#three)

4. [Aggregate Functions](#four)

5. [Sorting and Grouping](#five)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(odbc)
library(DBI)
library(tidyverse)
library(RSQLite)
```

```{r include=FALSE}
films = read_excel("/Volumes/Vandy Main/Data Camp/SQL/Intermediate SQL data/films.xlsx")
people = read_excel("/Volumes/Vandy Main/Data Camp/SQL/Intermediate SQL data/people.xlsx")
reviews = read_excel("/Volumes/Vandy Main/Data Camp/SQL/Intermediate SQL data/reviews.xlsx")
```


---


```{r include=FALSE}
con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = ":memory:")

#store sample data in database
dbWriteTable(conn = con, 
             name = "films",
             value = films)

dbWriteTable(conn = con, 
             name = "people",
             value = people)

dbWriteTable(conn = con, 
             name = "reviews",
             value = reviews)

#remove the local data from the environment
rm(films, people, reviews)
```


---

<a name="one"><a/>

## View the tables

---

#### View the `films` table.

```{sql echo=TRUE, connection=con, output.var="films"}
SELECT *
FROM films
LIMIT 10
```
```{R, echo=FALSE}
print(films)
```


---

#### View the `people` table. 

```{sql echo=TRUE, connection=con, output.var="people"}
SELECT *
FROM people
LIMIT 10
```
```{R, echo=FALSE}
print(people)
```


---

#### View the `reviews` table. 

```{sql echo=TRUE, connection=con, output.var="reviews"}
SELECT *
FROM reviews
LIMIT 10
```
```{R, echo=FALSE}
print(reviews)
```

---

<a name="two"><a/>

## Selecting and Counting Data

---

#### Count the number of observations in the `film_id` column of the `reviews` table.

```{sql echo=TRUE, connection=con, output.var="s1"}
SELECT COUNT(film_id) AS count_film_id
FROM reviews;
```
```{R, echo=FALSE}
print(s1)
```


---

#### Count the number of records in the `people` table.

```{sql echo=TRUE, connection=con, output.var="s2"}
SELECT COUNT(ID) AS count_records
FROM people;
```
```{R, echo=FALSE}
print(s2)
```


---

#### Count the number of birthdates in the `people` table.

```{sql echo=TRUE, connection=con, output.var="s3"}
SELECT COUNT(birthdate) AS count_birthdate
FROM people;
```
```{R, echo=FALSE}
print(s3)
```


---

#### Count the languages and countries represented in the films table.

```{sql echo=TRUE, connection=con, output.var="s4"}
SELECT COUNT(language) AS count_language, COUNT(country) AS count_country
FROM films;
```
```{R, echo=FALSE}
print(s4)
```


---

#### Return the unique countries from the `films` table.

```{sql echo=TRUE, connection=con, output.var="s5"}
SELECT DISTINCT (country)
FROM films
LIMIT(10)
```
```{R, echo=FALSE}
print(s5)
```


---

#### Count the distinct countries from the films table.

```{sql echo=TRUE, connection=con, output.var="s6"}
SELECT COUNT(DISTINCT(country)) AS count_distinct_countries
FROM films;
```
```{R, echo=FALSE}
print(s6)
```

---

<a name="three"><a/>

## Filtering Numbers

---

#### Select `film_ids` and `imdb_score` with an imdb_score over 7.0 from the `reviews` table.

```{sql echo=TRUE, connection=con, output.var="F1"}
SELECT film_id, imdb_score
FROM reviews
WHERE imdb_score > 7
LIMIT 10;
```
```{R, echo=FALSE}
print(F1)
```


---

#### Select the `film_id` and `facebook_likes` of the first ten records with less than 1000 likes from the `reviews` table.

```{sql echo=TRUE, connection=con, output.var="F2"}
SELECT film_id, facebook_likes
FROM reviews
WHERE facebook_likes < 1000
LIMIT 10;
```
```{R, echo=FALSE}
print(F2)
```


---

#### Count the records with at least 100,000 votes from the `reviews` table.

```{sql echo=TRUE, connection=con, output.var="F3"}
SELECT COUNT(num_votes) AS films_over_100k_votes
FROM reviews
WHERE num_votes >= 100000;
```
```{R, echo=FALSE}
print(F3)
```


---

#### Count the Spanish-language films from the `films` table.

```{sql echo=TRUE, connection=con, output.var="F4"}
SELECT COUNT(language) AS count_spanish
FROM films
WHERE language = 'Spanish';
```
```{R, echo=FALSE}
print(F4)
```


---

#### Select the `title` and `release_year` for all German-language films released before 2000 from the `films` table.

```{sql echo=TRUE, connection=con, output.var="F5"}
SELECT title, release_year
FROM films
WHERE language = 'German' AND release_year < 2000;
```
```{R, echo=FALSE}
print(F5)
```


---

#### Select all records for German-language films released after 2000 and before 2010 from the `films` table.

```{sql echo=TRUE, connection=con, output.var="F6"}
SELECT *
FROM films
WHERE language = 'German' 
		AND release_year > 2000
		AND release_year < 2010;
```
```{R, echo=FALSE}
print(F6)
```


---

#### Find the `title` and `year` of films from the year 1990 or 1999 from the `films` table.

- Add a filter to see only English or Spanish-language films.

- Filter films with more than $2,000,000 gross.

```{sql echo=TRUE, connection=con, output.var="F7"}
SELECT title, release_year
FROM films
WHERE (release_year = 1990 OR release_year = 1999)
	AND (language = 'Spanish' OR language = 'English')
	AND (gross > 2000000)
LIMIT 10;

```
```{R, echo=FALSE}
print(F7)
```


---

#### Select the `title` and `release_year` for films released between 1990 and 2000 from the `films` table.

- Narrow down the query to films with budgets > $100 million.

- Restrict the query to only Spanish or French language films.

```{sql echo=TRUE, connection=con, output.var="F8"}
SELECT title, release_year
FROM films
WHERE (release_year BETWEEN 1990 AND 2000)
	AND (budget > 100000000)
	AND (language = 'Spanish' OR language = 'French');
```
```{R, echo=FALSE}
print(F8)
```


---

#### Select the names that start with B from the `people` table.

```{sql echo=TRUE, connection=con, output.var="F9"}
SELECT name
FROM people
WHERE name LIKE 'B%'
LIMIT 10;
```
```{R, echo=FALSE}
print(F9)
```


---

#### Select the names that have r as the second letter from the `people` table.

```{sql echo=TRUE, connection=con, output.var="F10"}
SELECT name
FROM people
WHERE name LIKE '_r%'
LIMIT 10;
```
```{R, echo=FALSE}
print(F10)
```


---

#### Select names that don't start with A.

```{sql echo=TRUE, connection=con, output.var="F11"}
SELECT name
FROM people
WHERE name NOT LIKE 'A%'
LIMIT 10;
```
```{R, echo=FALSE}
print(F11)
```


---

#### Find the `title` and `release_year` for all films over two hours in length released in 1990 and 2000 in the `films` table.

```{sql echo=TRUE, connection=con, output.var="F12"}
SELECT title, release_year
FROM films
WHERE duration > 120
	AND release_year IN(1990, 2000)
LIMIT 10;
```
```{R, echo=FALSE}
print(F12)
```


---

#### Find the `title` and `language` of all films in English, Spanish, and French in the `films` table.

```{sql echo=TRUE, connection=con, output.var="F13"}
SELECT title, language
FROM films
WHERE language IN('English', 'Spanish', 'French')
LIMIT 10;
```
```{R, echo=FALSE}
print(F13)
```


---

#### Find the `title`, `certification`, and `language` of all films certified NC-17 or R that are in English, Italian, or Greek in the `films` table.

```{sql echo=TRUE, connection=con, output.var="F14"}
SELECT title, certification, language
FROM films
WHERE language IN('English', 'Italian', 'Greek')
	AND certification IN('NC-17', 'R')
LIMIT 10;
```
```{R, echo=FALSE}
print(F14)
```


---

#### Count the unique titles from the `films` table.

- Filter for release years between 1990 and 1999.

- Filter to English-language films.

- Narrow it down to G, PG, and PG-13 certifications.

```{sql echo=TRUE, connection=con, output.var="F15"}
SELECT COUNT(DISTINCT(title)) AS nineties_english_films_for_teens
FROM films
WHERE release_year BETWEEN 1990 AND 1999
	AND language = 'English'
	AND certification IN ('G', 'PG', 'PG-13');
```
```{R, echo=FALSE}
print(F15)
```


---

#### List all film titles with missing budgets from the `films` table.

```{sql echo=TRUE, connection=con, output.var="F16"}
SELECT title AS no_budget_info
FROM films
WHERE budget IS NULL
LIMIT 10;
```
```{R, echo=FALSE}
print(F16)
```


---

#### Count the number of films we have language data for.

```{sql echo=TRUE, connection=con, output.var="F17"}
SELECT COUNT(*) AS count_language_known
FROM films
WHERE language IS NOT NULL; 
```
```{R, echo=FALSE}
print(F17)
```


---

<a name="four"><a/>

# Aggregate Functions

---

#### Query the sum of film durations from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A1"}
SELECT SUM(duration) AS total_duration
FROM films;
```
```{R, echo=FALSE}
print(A1)
```


---

#### Calculate the average duration of all films from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A2"}
SELECT AVG(duration) AS average_duration
FROM films;
```
```{R, echo=FALSE}
print(A2)
```


---

#### Find the latest `release_year` from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A2"}
SELECT MAX(release_year) AS latest_year
FROM films;
```
```{R, echo=FALSE}
print(A2)
```


---

#### Find the duration of the shortest film from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A3"}
SELECT MIN(duration) AS shortest_film
FROM films;
```
```{R, echo=FALSE}
print(A3)
```


---

#### Calculate the sum of gross from the year 2000 or later from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A4"}
SELECT SUM(gross) AS total_gross
FROM films
WHERE release_year >= 2000;
```
```{R, echo=FALSE}
print(A4)
```


---

#### Calculate the average gross of films that start with A from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A5"}
SELECT AVG(gross) AS avg_gross_A
FROM films
WHERE title LIKE 'A%';
```
```{R, echo=FALSE}
print(A5)
```

---

#### Calculate the lowest gross film in 1994 from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A6"}
SELECT MIN(gross) AS lowest_gross
FROM films
WHERE release_year = 1994;
```
```{R, echo=FALSE}
print(A6)
```

---

#### Calculate the highest gross film released between 2000-2012 from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A7"}
SELECT MAX(gross) AS highest_gross
FROM films
WHERE release_year BETWEEN 2000 AND 2012;
```
```{R, echo=FALSE}
print(A7)
```

---

#### Round the average number of `facebook_likes` to one decimal place from the `reviews` table.

```{sql echo=TRUE, connection=con, output.var="A8"}
SELECT ROUND(AVG(facebook_likes), 1) AS avg_facebook_likes
FROM reviews;
```
```{R, echo=FALSE}
print(A8)
```

---

#### Calculate the average budget rounded to the thousands from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A9"}
SELECT ROUND(AVG(budget), -3) AS avg_budget_thousands
FROM films;
```
```{R, echo=FALSE}
print(A9)
```

---

#### Calculate the `title` and `duration_hours` from `films`.

```{sql echo=TRUE, connection=con, output.var="A10"}
SELECT title, duration/60 AS duration_hours
FROM films
LIMIT 10;
```
```{R, echo=FALSE}
print(A10)
```

---

#### Calculate the percentage of people who are no longer alive from the `people` table.

```{sql echo=TRUE, connection=con, output.var="A11"}
SELECT (COUNT(deathdate)*100.00)/COUNT(*) AS percentage_dead
FROM people;
```
```{R, echo=FALSE}
print(A11)
```

---

#### Find the number of decades in the `films` table.

```{sql echo=TRUE, connection=con, output.var="A12"}
SELECT (MAX(release_year) - MIN(release_year)) / 10 AS number_of_decades
FROM films;
```
```{R, echo=FALSE}
print(A12)
```

---

#### Select the `title` and round `duration_hours` to two decimal places from the `films` table.

```{sql echo=TRUE, connection=con, output.var="A13"}
SELECT title, ROUND(duration/60.0, 2) AS duration_hours
FROM films
LIMIT 10;
```
```{R, echo=FALSE}
print(A13)
```

---

<a name="five"><a/>

# Sorting and Grouping.

---

#### Select `name` from `people` and sort alphabetically.

```{sql echo=TRUE, connection=con, output.var="SG1"}
SELECT name
FROM people 
ORDER BY name
LIMIT 10;
```
```{R, echo=FALSE}
print(SG1)
```

---

#### Select the `title` and `duration` from longest to shortest film from the `films` table.

```{sql echo=TRUE, connection=con, output.var="SG2"}
SELECT title, duration
FROM films
ORDER BY duration DESC
LIMIT 10;
```
```{R, echo=FALSE}
print(SG2)
```

---

#### Select the `release year`, `duration`, and `title` sorted by release year and duration from the `films` table.

```{sql echo=TRUE, connection=con, output.var="SG3"}
SELECT release_year, duration, title
FROM films
ORDER BY release_year, duration
LIMIT 10;
```
```{R, echo=FALSE}
print(SG3)
```

---

#### Select the `certification`, `release year`, and `title` sorted by certification and release year from the `films` table.

```{sql echo=TRUE, connection=con, output.var="SG4"}
SELECT certification, release_year, title
FROM films
ORDER BY certification, release_year
LIMIT 10;
```
```{R, echo=FALSE}
print(SG4)
```

---

#### Find the `release_year` and `film_count` of each year from the `films` table.

```{sql echo=TRUE, connection=con, output.var="SG5"}
SELECT release_year, COUNT(title) AS film_count
FROM films
GROUP BY release_year
LIMIT 10;
```
```{R, echo=FALSE}
print(SG5)
```

---

#### Find the `release_year` and `average duration` of films for each year from the `films` table.

```{sql echo=TRUE, connection=con, output.var="SG6"}
SELECT release_year, AVG(duration) AS avg_duration
FROM films
GROUP BY release_year
LIMIT 10;
```
```{R, echo=FALSE}
print(SG6)
```

---

#### Find the `release_year`, `country`, and `max_budget` from the `films table`.

- Group and order by `release_year` and `country`.

```{sql echo=TRUE, connection=con, output.var="SG7"}
SELECT release_year, country, MAX(budget) AS max_budget
FROM films
GROUP BY release_year, country
ORDER BY release_year, country
LIMIT 10;
```
```{R, echo=FALSE}
print(SG7)
```

---

#### Which `release_year` had the most language diversity?

```{sql echo=TRUE, connection=con, output.var="SG8"}
SELECT release_year, COUNT(DISTINCT(language)) AS lang
FROM films 
GROUP BY release_year
ORDER BY lang DESC
LIMIT 1;
```
```{R, echo=FALSE}
print(SG8)
```

---

#### Select the `country` and distinct count of certification as `certification_count` from the films table.

- Group by country.

- Filter results to countries with more than 10 different certifications.

```{sql echo=TRUE, connection=con, output.var="SG9"}
SELECT country, COUNT(DISTINCT(certification)) AS certification_count
FROM films 
GROUP BY country
HAVING COUNT(DISTINCT(certification)) > 10;
```
```{R, echo=FALSE}
print(SG9)
```

---

#### Select the `country` and `average_budget` from the `films` table.

- Group by `country`.

- Filter to countries with an `average_budget` of more than one billion.

- Order by descending order of the aggregated budget.

```{sql echo=TRUE, connection=con, output.var="SG9"}
SELECT country, AVG(budget) AS average_budget
FROM films 
GROUP BY country
HAVING AVG(budget) > 1000000000
ORDER BY average_budget DESC;
```
```{R, echo=FALSE}
print(SG9)
```

---

#### Select the average budget for films released after 1990 from the `films` table.

- Group by year.

```{sql echo=TRUE, connection=con, output.var="SG10"}
SELECT release_year, AVG(budget) AS avg_budget
FROM films 
GROUP BY release_year
HAVING release_year > 1990
ORDER BY release_year DESC
LIMIT 10;
```
```{R, echo=FALSE}
print(SG10)
```

---

#### Select `release year`, `average budget`, and `average gross` from the `films` table and group by release year.

- Limit results to films released after the year 1990.

- Limit results to films with an average budget of more than 60 million.

- Order the results from the highest average gross and limit to one.

```{sql echo=TRUE, connection=con, output.var="SG11"}
SELECT release_year, AVG(budget) AS avg_budget, AVG(gross) AS avg_gross
FROM films
WHERE release_year > 1990
GROUP BY release_year
HAVING AVG(budget) > 60000000
ORDER BY avg_gross DESC 
LIMIT 1;
```
```{R, echo=FALSE}
print(SG11)
```













